---
title: "p8105_hw3_yh3824"
author: "ruby"
date: "2025-10-02"
output: html_document
---

```{r include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
library(p8105.datasets)
data("instacart")

instacart
```

The `instacart` dataset containts `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each row represents a product from an order. There are `r n_distinct(instacart$order_id)` unique order_ids. 

```{r}
aislecnt = instacart |> 
  count(aisle, name = "n_obs") |> 
  arrange(desc(n_obs))
```

There are 134 aisles and the aisles with the most items ordered are `fresh vegetables` with 150,609 items, `fresh fruits` with 150,473 items, `packaged vegetable fruits` with 78,493 items. 

### Plot
```{r}
aislecnt |> 
  filter(n_obs > 10000) |> 
  ggplot(aes(x = reorder(aisle, n_obs), y = n_obs)) +
  geom_col(width = 0.7) +
  scale_y_continuous(breaks = seq(10000, 150700, by = 20000)) +
  theme (
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  )
```

### Table with 3 most popular items in each of the aisles 
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(count = n()) |> 
  arrange(desc(count)) |> 
  slice_head(n=3) |> 
  knitr::kable(digits = 1)
  
```

### Table showing the mean hour of the day at which Pink Lady Apples + Coffee Ice Cream are consumed 
```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  mutate(order_dow = recode(order_dow,
                            `0` = "Sunday", `1` = "Monday", `2` = "Tuesday",
                            `3` = "Wednesday", `4` = "Thursday",
                            `5` = "Friday", `6` = "Saturday")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 1)
```


## Problem 2
```{r}
zipcode_df = 
  read_csv("zillow_data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  mutate(zip_code = as.integer(zip_code)) |> 
  select(-state_fips, - county_code, -county_fips, -file_date) |> 
  distinct(zip_code, .keep_all = TRUE)

zip_codes_df = 
  read_csv("zillow_data/Zip Codes.csv") |> 
  janitor::clean_names() |> 
  filter(
    !(zip_code == 10463 & county == "New York"),
    !(zip_code == 11201 & county == "New York"))  |> 
  mutate(
    borough = case_match(
      county,
      "Bronx" ~ "Bronx",
      "Kings" ~ "Brooklyn",
      "New York" ~ "Manhattan",
      "Queens" ~ "Queens",
      "Richmond" ~ "Staten Island"
    )
  ) |> 
  select(zip_code, borough, neighborhood)




```

import and clean zori data
```{r}
zori_df =
  read_csv("zillow_data/Zip_Zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  janitor::clean_names() |> 
  mutate(zip_code = as.integer(region_name)) |> 
  pivot_longer(x2015_01_31:x2024_08_31, 
               names_to = "date", 
               values_to = "zori",
               names_prefix = "x") |> 
  separate(date, into = c("year", "month", "day"), sep = "_", convert = TRUE) |> 
  rename(county = county_name) |> 
  mutate(county = str_remove(county, 
                                  regex(" county$", ignore_case = TRUE))) |> 
  select(county, zip_code, year, month, zori) 

nyc_zillow_df = 
  read_csv("zillow_data/Zip_Zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  pivot_longer(
    -(RegionID:CountyName),
    names_to = "month",
    values_to = "price") |> 
  janitor::clean_names() |> 
  rename(zip_code = region_name) |> 
  mutate(
    month = as_date(month),
    year = floor_date(month, unit = "year"),
    zip_code = as.numeric(zip_code)) |> 
  select(-county_name)

nyc_price_df = 
  left_join(nyc_zillow_df, zip_codes_df, by = "zip_code")
```

### How many zip codes are observed 116 times, how many are observed fewer than 10 times, why are some zip codes observed rarely and others observed in each month 
```{r}
nyc_price_df |> 
  group_by(zip_code) |> 
  summarize(count = n()) |> 
  filter(count > 115) |> 
  knitr::kable(digits = 1)
```

149 zip codes are observed 116 times. All 

### reader friendly table showing avg rental price in each borough and YEAR 
```{r}
nyc_price_df |> 
  group_by(borough, year) |> 
  summarize(avg_price = mean(price, na.rm = TRUE)) |> 
  pivot_wider(
    names_from = year, 
    values_from = avg_price
  ) |> 
  knitr::kable(digits = 1)
```

### Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot. 
```{r}
price_zip_allyears = 
nyc_price_df |> 
  group_by(borough, zip_code, year) |> 
  summarize(avg_price = mean(price, na.rm = TRUE)) |> 
  ggplot(aes(x = borough, y = avg_price)) +
  geom_boxplot() +
  facet_wrap(~year)
```

### compute avg rental price within each zip code over each month in 2023 + make a plot
```{r}
price_zip_2023 =
nyc_price_df |> 
  filter(year == "2023-01-01") |> 
  group_by(zip_code, month, borough) |> 
  summarize(avg_price = mean(price, na.rm = TRUE)) |> 
  ggplot(aes(x = borough, y = avg_price)) +
  geom_boxplot() 
```

### combine 2 previous plots into a single graphic and export into a results folder in repo
```{r}
price_zip_allyears / price_zip_2023
```


## Problem 3

Your final dataset should include all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind)

```{r}
demo_df = 
  read_csv("nhanes_covar.csv", skip = 4, na = "NA") |> 
  janitor::clean_names() |> 
  filter(age > 20) |> 
  na.omit() |> 
  mutate(sex = ifelse(sex == 1, "male", "female"), 
         education = case_when(
    education == 1 ~ "Less than high school",
    education == 2 ~ "High school equivalent",
    education == 3 ~ "More than high school"))

accel_df = 
  read_csv("nhanes_accel.csv", na = "NA") |> 
  janitor::clean_names() |> 
  na.omit()


combine_df = 
  left_join(demo_df, accel_df, by = "seqn")
```

i do not think i need a pivot longer in this question!!
pivot_longer(
    cols = starts_with("min"),
    names_to = "min_num",
    values_to = "min_value",
    names_prefix = "min"
    
    
### produce a reader-friendly table for the number of men and woemn in each education category, create a visualization of the age distributions for men and women in each education category 


```{r}
combine_df |> 
  group_by(sex, education) |> 
  summarize(count = n()) |> 
  pivot_wider(
    names_from = education,
    values_from = count
  ) |> 
  knitr::kable(digits = 1)

combine_df |> 
  ggplot(aes(x = age)) +
  geom_histogram() +
  facet_grid(sex ~ education)
```

### aggregate across minutes to create a total activity variable for each participant, plot y = total activity, x = age, should compare men to women and have separate panels for each education level, include a trend line or smooth to illustrate differences 
```{r}

```

### make a 3 panel plot that shows the 24-hr activity time courses for each education level and use color to indicate sex, describe in words any patterns or conclusions you can make based on this graph including a smooth trends may help