---
title: "p8105_hw3_yh3824"
author: "ruby"
date: "2025-10-02"
output: html_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
library(p8105.datasets)
data("instacart")

instacart
```

The `instacart` dataset containts `r nrow(instacart)` observations and `r ncol(instacart)` variables. Each row represents a product from an order. There are `r n_distinct(instacart$order_id)` unique order_ids. 

```{r}
aislecnt = instacart |> 
  count(aisle, name = "n_obs") |> 
  arrange(desc(n_obs))
```

There are 134 aisles and the aisles with the most items ordered are `fresh vegetables` with 150,609 items, `fresh fruits` with 150,473 items, `packaged vegetable fruits` with 78,493 items. 

### Plot
```{r}
aislecnt |> 
  filter(n_obs > 10000) |> 
  ggplot(aes(x = n_obs, y = reorder(aisle, n_obs))) +
  geom_col(width = 0.5) +
  scale_x_continuous(breaks = seq(10000, 150700, by = 20000)) +
  theme (
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12)
  )
```

### Table with 3 most popular items in each of the aisles 
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  group_by(aisle, product_name) |> 
  summarize(count = n()) |> 
  arrange(desc(count)) |> 
  slice_head(n=3) |> 
  knitr::kable(digits = 1)
  
```

### Table showing the mean hour of the day at which Pink Lady Apples + Coffee Ice Cream are consumed 
```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  mutate(order_dow = recode(order_dow,
                            `0` = "Sunday", `1` = "Monday", `2` = "Tuesday",
                            `3` = "Wednesday", `4` = "Thursday",
                            `5` = "Friday", `6` = "Saturday")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 1)
```


